Draft genomes of the three northern hemisphere blue mussel lineages: North and South European \emph{Mytilus edulis} and Mediterranean \emph{Mytilus galloprovincialis}.

### 2022-06-30 manuscript/ms_genomes.tex
repeats (mainly)

We used \texttt{AGOUTI} (\url{https://github.com/svm-zhang/AGOUTI}; commit a7e65d6; \cite{Zhang2016a}) to improve scaffolding by using paired end RNA-seq reads.
For each species, a different set of published transcriptomes were used (Supplementary Table [[...]] for accessions).
\texttt{AGOUTI} require a gene prediction as input in addition to RNA-seq reads.
We used \texttt{Augustus} (v3.3.3; \cite{Stanke2008}) to produce an intermediate annotation for each assembly using the \textit{Caenorhabditis} model.
RNA-seq reads were first cleaned using \texttt{rcorrector} (v1.0.4; \cite{Song2015}) and \texttt{trimgalore} (v0.6.6; \url{https://www.bioinformatics.babraham.ac.uk/projects/trim_galore}; \texttt{--quality 20 \\--stringency 1 -e 0.1 --length 70}).
RNA-seq reads were mapped independently using \texttt{bwa-mem2} (v2.2.1; \cite{VasimuddinMd2019}), and then merged as a common bam file for each assembly with \texttt{samtools} (v1.12; \cite{Li2009}).
Finally, the \texttt{AGOUTI} scaffolding pipeline was run using the gene prediction and mapped RNA-seq reads (\texttt{-minMQ 20 -maxFracMM 0.05}; using \texttt{python} v2.7.15 and \texttt{samtools} v1.10 for compatibility).

@@

To obtain an annotation for each assembly, we used both database and RNA-seq information.
We used \texttt{Braker2} (v2.1.6; \cite{Bruna2021}) as an initial step, using both a protein database and RNA-seq reads (preprocessed in \cref{draft_asm}).
To build the protein database, we used all \textit{Mollusca} proteins (taxid 6447) from OrthoDB (v10.1; \cite{Kriventseva2019}).
To provide gene presence hints, we mapped all RNA-seq reads for each species using \texttt{HISAT2} (v2.2.1; \cite{Kim2019}).

We used the \texttt{Mantis} pipeline (\url{https://github.com/PedroMTQ/mantis}; commit c6cb597; \cite{Queiros2021}) to obtain consensus annotations of genes based on multiple databases.
Protein sequences for each assembly were built from the \texttt{Braker2} annotation using the \texttt{python} module \texttt{gff3tool} (v2).

@@

We used \texttt{OrthoFinder} (v2.5.2; \cite{Emms2015, Emms2019}) to find orthogroups and orthologue genes.
Species tree was inferred using the MSA method of \texttt{OrthoFinder} \parencite{Emms2018} and \texttt{IQTREE} (v2.2.0\_beta; \cite{Minh2020}) tree inference.

@@

We introduce here newly assembled genomes from three lineages of the \textit{Mytilus} species complex.
With a limited budget of $\sim$3000€ per genome, we managed to produce draft genomes of enough quality to be useful
in applications such as population genomics and genetics.
The method of 10X chromium pseudo-long reads combined with scaffolding using published data provided a quality comparable to published assemblies for \textit{Mytilus} species.

The pre-assembly k-mer analysis carried out using GenomeScope (using 21-mers) showed that, as expected, the genomes were highly heterozygous with values going from 3.49 to 8.77~\% (\cref{fig:genomescope}).
GenomeScope provide a bad fit and estimation in the case of the MeduEUS data due to a reduced coverage of the genome in that case, compared to the two other assemblies.

@@

@@

The species tree (\cref{fig:tree}) shows that the assemblies and the published genomes and transcriptomes are clustering as expected.
\texttt{OrthoFinder} \comment{provides}{provide the orthogroups somewhere} a resource of single-copy orthogroups across a large part the \textit{Mytilus} species complex.

OrthoFinder results:
OrthoFinder assigned 499004 genes (91.9\% of total) to 50543 orthogroups. Fifty percent of all 
genes were in orthogroups with 13 or more genes (G50 was 13) and were contained in the largest
11982 orthogroups (O50 was 11982). There were 1852 orthogroups with all species present and 
173 of these consisted entirely of single-copy genes.

Species tree using 1300 orthogroups with minimum of 81.8\% of species having single-copy genes in any orthogroup (value automatically selected by \texttt{OrthoFinder}).

\begin{figure}[h]
	\includegraphics[width=\linewidth]{figures/Fig3_tree.pdf}
	\caption{Species tree using 1300 orthogroups with minimum of 81.8\% of species having single-copy genes in any orthogroup.
	Color coding -- dark blue: \textit{M. edulis} Europe, light blue: \textit{M. edulis} America,
	red: \textit{M. galloprovincialis} Mediterranean Sea, yellow: \textit{M. galloprovincialis} Atlantic,
	green: \textit{M. trossulus}, pink: \textit{M. coruscus}, gray: \textit{M. californianus}.
	}
	\label{fig:tree}
\end{figure}

@@

Raw reads are available under BioProject PRJNA785550.
Assemblies are available under accessions JAKGDF000000000 (MgalMED), JAKGDG000000000 (MeduEUS), and JAKGDH000000000 (MeduEUN).
Annotations are available under the Zenodo archive [XXX].
The assembly pipeline is available at \url{https://github.com/alxsimon/assembly_10x}.
The OrthoFinder pipeline is available at [XXX].

@@

We thank [XXX].
AS was supported by the ANR grant [XXX] and by the National Institute of General Medical Sciences of the National Institutes of Health (grants NIH R01 GM108779 and R35 GM136290 to Dr Graham Coop).

@@

>>>>

We used \texttt{AGOUTI} (\url{https://github.com/svm-zhang/AGOUTI}; commit a7e65d6; \cite{Zhang2016a}) to improve scaffolding by using paired end RNA-seq reads.
For each species, a different set of published transcriptomes were used (See Supplementary File 1 for accession numbers).
\texttt{AGOUTI} require a gene prediction as input in addition to RNA-seq reads.
We used \texttt{Augustus} (v3.3.3; \cite{Stanke2008}) to produce an intermediate annotation for each assembly using the \textit{Caenorhabditis} model.
RNA-seq reads were first cleaned using \texttt{rcorrector} (v1.0.4; \cite{Song2015}) and \texttt{trimgalore} (v0.6.6; \url{https://www.bioinformatics.babraham.ac.uk/projects/trim_galore}; \texttt{--quality 20 \\--stringency 1 -e 0.1 --length 70}).
RNA-seq reads were mapped independently using \texttt{bwa-mem2} (v2.2.1; \cite{VasimuddinMd2019}), and then merged as a common bam file for each assembly with \texttt{samtools} (v1.12; \cite{Li2009}).
Finally, the \texttt{AGOUTI} scaffolding pipeline was run using the gene prediction and mapped RNA-seq reads (\texttt{-minMQ 20 -maxFracMM 0.05}; using \texttt{python} v2.7.15 and \texttt{samtools} v1.10 for compatibility).

@@

We used \texttt{Braker2} (v2.1.6; \cite{Bruna2021}) as to obtain structural annotations, using both a protein database and RNA-seq reads (preprocessed in \cref{draft_asm}).
To build the protein database, we used all \textit{Mollusca} proteins (taxid 6447) from OrthoDB (v10.1; \cite{Kriventseva2019}).
To provide gene presence hints, we mapped all RNA-seq reads for each species using \texttt{HISAT2} (v2.2.1; \cite{Kim2019}).

We used the \texttt{Mantis} pipeline (\url{https://github.com/PedroMTQ/mantis}; commit c6cb597; \cite{Queiros2021}) to obtain consensus functional annotations of genes based on multiple databases.
Protein sequences for each assembly were built from the \texttt{Braker2} annotation using the \texttt{python} module \texttt{gff3tool} (v2).

@@

We used \texttt{OrthoFinder} (v2.5.4; \cite{Emms2015, Emms2019}) to find orthogroups and orthologue genes.
Species tree was inferred using the MSA method of \texttt{OrthoFinder} \parencite{Emms2018} using the \texttt{MAFFT} aligner (v7.505; \cite{Katoh2013}), the \texttt{STRIDE} species tree rooting algorithm \parencite{Emms2017} and the \texttt{FastTree} software for tree inference (v2.1.11; \cite{Price2009}).

@@

\subsection{Assembly results}

We introduce here newly assembled genomes from three lineages of the \textit{Mytilus} species complex.
With a limited budget of $\sim3000$€ per genome, we managed to produce draft genomes of enough quality to be useful
in applications such as population genomics and genetics.
The method of 10X chromium pseudo-long reads combined with scaffolding using published data provided a quality comparable to published assemblies for \textit{Mytilus} species.

The pre-assembly k-mer analysis carried out using GenomeScope (using 21-mers) showed that, as expected, the genomes were highly heterozygous with values going from $3.49$ to $8.77$~\% (\cref{fig:genomescope}).
For the MeduEUS data, GenomeScope provides a fit of bad quality and an estimation probably off due to a reduced slightly lower sequencing depth compared to the two other assemblies. In that case, the heterozygous peak was not identifiable.

@@

Stats:
GC contents: 32.1\%, 34.6\%, 32.1\%

@@

The MeduEUS assembly suffered from an increased level of contamination compared to the other two assemblies.
Despite a broad contaminant filtering step, the Busco blob analysis (\cref{supfig:btk-blob-MeduEUS})
shows a second GC content peak centered around 40\% (above the ~32\% peaks of this and other genomes).

\subsection{Repeat contents}

More than half of each genome was identified as repeats and masked by \texttt{RepeatMasker}.
Repeats are estimated to amount to 57.22, 61.10 and 55.53\% of the bases of MgalMED, MeduEUS and MeduEUN assemblies respectively.
The majority of repeats are unclassified, followed by retroelements with a balanced contribution of LINEs and LTR elements
(detailed \texttt{RepeatMasker} results in \cref{suptab:repeats}).

\subsection{Annotations}

[[Here goes annotation stats]]

\subsection{Species tree}

OrthoFinder assigned 499004 genes (91.9\% of total) to 50543 orthogroups. Fifty percent of all 
genes were in orthogroups with 13 or more genes (G50 was 13) and were contained in the largest
11982 orthogroups (O50 was 11982). There were 1852 orthogroups with all species present and 
173 of these consisted entirely of single-copy genes.

The species tree was built using 1300 orthogroups with a minimum of 81.8\% of species having single-copy genes in any orthogroup (\cref{fig:tree}).
It shows that the assemblies and the published genomes and transcriptomes are clustering as expected.

\begin{figure}[h]
	\includegraphics[width=\linewidth]{figures/Fig3_tree.pdf}
	\caption{Species tree using 1300 orthogroups with a minimum of 81.8\% of species having single-copy genes in any orthogroup.
	Color coding -- dark blue: \textit{M. edulis} Europe, light blue: \textit{M. edulis} America,
	red: \textit{M. galloprovincialis} Mediterranean Sea, yellow: \textit{M. galloprovincialis} Atlantic,
	green: \textit{M. trossulus}, pink: \textit{M. coruscus}, gray: \textit{M. californianus}.
	}
	\label{fig:tree}
\end{figure}

@@

Raw data are available under BioProject PRJNA785550.
Assemblies are available under accessions JAKGDF000000000 (MgalMED), JAKGDG000000000 (MeduEUS), and JAKGDH000000000 (MeduEUN).
The assembly pipeline is available at \url{https://github.com/alxsimon/assembly_10x}.
Annotations and the OrthoFinder pipeline and results are available on the Zenodo archive [[XXX]].

@@

We thank [[XXX]].
We thank IFB core cluster for computational resources [[...]]
We thank Tiago Horri for providing his annotation of the American \textit{Mytilus edulis} genome (GCA\_019925275.1).
AS was supported by the ANR grant [[XXX]] and by the National Institute of General Medical Sciences of the National Institutes of Health (grants NIH R01 GM108779 and R35 GM136290 to Dr Graham Coop).

@@

% Table contains discrepancies of the sums but this is what RepeatMasker outputs... So I don't know...
\include{figures/Suptab_repeats.tex}
